---
title: "p8105_hw1_cm3341"
author: "Carolina Montes Garcia"
date: "2024-09-17"
output: 
  github_document:
    toc: true 
editor_options: 
  chunk_output_type: console
---

Load tidyverse for its functions, tidy syntax, tibbles, and because it includes ggplot2

```{r}
library(tidyverse)
```


## Problem 1
The 'penguins' dataset includes 344 entries of bill measurements, flipper measurements, body mass in grams, sex, and the year of measurement for three different species of penguin - Adelie, Chinstrap, and Gentoo. 

```{r}
data("penguins", package = "palmerpenguins")

dataset_size = penguins %>%
  summarise(rows = n(), columns = ncol(penguins))

dataset_size
```


```{r}
mean_flipper_length = penguins %>%
  summarise(mean_flipper_length_mm = mean(flipper_length_mm, na.rm = TRUE))

mean_flipper_length
```


```{r}

```

### Scatterplot of flipper length vs. bill length
```{r}
ggplot(
  penguins, aes(
    x = bill_length_mm, 
    y = flipper_length_mm, 
    color = species)) +
  geom_point() +
  labs(
    title = "Flipper Length vs Bill Length", x = "Bill Length (mm)", y = "Flipper Length (mm)") +
  theme_minimal()

ggsave("flipper_vs_bill_length.png")
```


## Problem 2

Set seed for reproducibility purposes
```{r}
set.seed(123)
```

Create dataframe
```{r}
example_data = data.frame(
  normal_sample = rnorm(10),
  greater_than_zero = rnorm(10) > 0,
  char_vector = letters[1:10],
  factor_vector = factor(sample(c("fun level", "ugly level", "evil level"), 10, replace = TRUE))
)
```

Mean of each dataframe variable
```{r}


```

