---
title: "p8105_hw1_cm3341"
author: "Carolina Montes Garcia"
date: "2024-09-17"
output: 
  github_document:
    toc: true 
editor_options: 
  chunk_output_type: console
---

Load tidyverse for its functions, tidy syntax and operators, tibbles, and because it includes ggplot2.

```{r}
library(tidyverse)
```


## Problem 1
The 'penguins' dataset includes 344 entries of bill measurements, flipper measurements, body mass in grams, sex, and the year of measurement for three different species of penguin - Adelie, Chinstrap, and Gentoo. It also lists the island for each penguin. 


```{r}
data("penguins", package = "palmerpenguins")

dataset_size = penguins %>%
  summarise(rows = n(), columns = ncol(penguins))

dataset_size
```

The data set has `r ncol(penguins)` columns


```{r}
mean_flipper_length = penguins %>%
  summarise(mean_flipper_length_mm = mean(flipper_length_mm, na.rm = TRUE))

mean_flipper_length
```


```{r}

```

Scatterplot of flipper length vs. bill length
```{r message = FALSE, warning=FALSE}
ggplot(
  penguins, aes(
    x = bill_length_mm, 
    y = flipper_length_mm, 
    color = species)) +
  geom_point() +
  labs(
    title = "Flipper Length vs Bill Length", x = "Bill Length (mm)", y = "Flipper Length (mm)") +
  theme_minimal()

ggsave("flipper_vs_bill_length.png")
```


## Problem 2

Setting a seed for sampling consistency and reproducibility
```{r}
set.seed(123)
```

Create dataframe
```{r}

example_data = data.frame(
  norm_samp = rnorm(10),
  norm_samp_pos = rnorm(10) > 0,
  char_vector = letters[1:10],
  factor_vector = factor(sample(c("fun level", "ugly level", "evil level"), 10, replace = TRUE))
)

```

Mean of each dataframe variable
```{r}

mean_samp1 = mean(pull(example_data, norm_samp))
mean_samp2 = mean(pull(example_data, norm_samp_pos))
mean_samp3 = mean(pull(example_data, char_vector))
mean_samp4 = mean(pull(example_data, factor_vector))

```

*The mean of the random sample is `r mean(pull(example_data, norm_samp))`

*The mean of the logical vector is `r mean(pull(example_data, norm_samp_pos))`, which is strange at first glance because the entries in the logical vector look like characters. However, R seems to treat *TRUE* as a *1* and *FALSE* as a *0* and so it calculates the mean based on the proportion of *TRUE* entries in the vector. 

*The mean of the character vector is `r mean(pull(example_data, char_vector))`, which makes sense because one cannot calculate the mean of a categorical variable. 

*The mean of the character vector is `r mean(pull(example_data, factor_vector))`, which makes sense because one cannot calculate the mean of a categorical variable. 

```{r, eval = FALSE}
example_data[, "norm_samp_pos"] <- as.numeric(df[, "norm_samp_pos"])


num_norm_samp_pos = as.numeric(norm_samp_pos)

```

