---
title: "p8105_hw1_cm3341"
author: "Carolina Montes Garcia"
date: "2024-09-17"
output: 
  github_document:
    toc: true 
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
install.packages("palmerpenguins")

```

##Load packages that I need for tidyverse functions and plotting
```{r}
library(tidyverse)
library(ggplot2)
```


## Problem 1
The 'penguins' dataset includes bill length and flipper length of three different species of penguin - Adelie, Chinstrap, and Gentoo. 

    the data in this dataset, including names / values of important variables
    the size of the dataset (using nrow and ncol)
    the mean flipper length



```{r}
data("penguins", package = "palmerpenguins")


```

## Scatterplot of flipper length vs. bill length
```{r}
ggplot(
  penguins, aes(
    x = bill_length_mm, 
    y = flipper_length_mm, 
    color = species)) +
  geom_point() +
  labs(
    title = "Flipper Length vs Bill Length", x = "Bill Length (mm)", y = "Flipper Length (mm)") +
  theme_minimal()

ggsave("flipper_vs_bill_length.png")
```


## Problem 2

```{r}

```

